A Meta-analysis of Longevity Estimates of Mosquito Vectors of Disease: Additional Methods

### 2019-06-17 writeUp/supplementary_materials_methods.tex
Updated SOM

\section{Detinova parity analysis}
\cite{massey2016global} provide a global database for bionomic quantities for the dominant vector species of malaria. In this database, there are

\cite{detinova1962age} provided an alternative approach to estimate the lifespan of individual specimens, based on sampling the parous rate in a population (the proportion of adult females that have laid eggs) by separating parous from nulliparous females. In what follows, we follow the derivation given by \cite{davidson1954estimation}, to determine an expression for the proportion of nulliparous mosquitoes in terms of the gonotrophic cycle length $G$, here assumed the same for each cycle subsequent to an adult emerging. If a constant daily survival rate $p$ is assumed, the proportion of emerging adults surviving to become parous is $p^G$. If the rate of recruitment into the adult classes is constant over time, the proportion of nulliparous mosquitoes is,
%
\begin{equation}\label{eq:detinova}
\begin{aligned}
\mathbb{n} &= \frac{p^G}{p^G + p^{2G} + p^{3G} + \cdots}\\
&= p^G / \frac{p^G}{1-p^G}\\
&= 1-p^G.
\end{aligned}
\end{equation}
%
The daily survival rate can hence be calculated by inverting eq. (\ref{eq:detinova}): $p = \sqrt[G]{\mathcal{P}}$, where $\mathcal{P}=1-\mathbb{n}$ is the proportion parous sampled from the population.

>>>>

\section{Detinova parity analysis}
\cite{detinova1962age} provided an alternative approach to estimate the lifespan of individual specimens, based on sampling the parous rate in a population (the proportion of adult females that have laid eggs) using dissection to determine parity of each specimen. In what follows, we follow the derivation given by \cite{davidson1954estimation}, to determine an expression for the proportion of nulliparous mosquitoes in terms of the gonotrophic cycle length $G$, here assumed the same for each cycle subsequent to an adult emerging. If a constant daily survival rate $p$ is assumed, the proportion of emerging adults surviving to become parous is $p^G$. If the rate of recruitment into the adult classes is constant over time, the proportion of nulliparous mosquitoes is,
%
\begin{equation}\label{eq:detinova}
\begin{aligned}
\mathbb{n} &= \frac{p^G}{p^G + p^{2G} + p^{3G} + \cdots}\\
&= p^G / \frac{p^G}{1-p^G}\\
&= 1-p^G.
\end{aligned}
\end{equation}
%
The daily survival rate can hence be calculated by inverting eq. (\ref{eq:detinova}): $p = \sqrt[G]{\mathcal{P}}$, where $\mathcal{P}=1-\mathbb{n}$ is the proportion parous sampled from the population.

\cite{massey2016global} provide a global database of bionomic quantities for the dominant vector species of malaria. This database contains the number of parous and nulliparous samples in studies where Detinova's dissection approach was applied to field caught speciments ($n=1490$ non-missing observations) across Africa ($n=778$), Asia ($n=549$) and the Americas ($n=163$).

We estimated lifespan at the species and complex / subgroup level (hereafter denoted as \textit{s.l.}) using a Bayesian hierarchical framework. The number of parous individuals $X_i^{\{s\}}$ in sample $i$ (consisting of species $s$) was modelled by a binomial sampling distribution,
%
\begin{equation}
X_i^{\{s\}} \sim \mathcal{B}(N, \mathcal{P}_i^{\{s\}}),
\end{equation}
%
where $N$ is the sample size and $\mathcal{P}_i^{\{s\}}$ is the proportion parous in sample $i$ for species $s$. A hierarchical prior structure was used for the parous proportion in each sample,
%
\begin{equation}
\mathcal{P}_i^{\{s\}}\sim \text{beta}\left(\phi^{\{s\}} \kappa^{\{s\}}, (1-\phi^{\{s\}})\kappa^{\{s\}}\right),
\end{equation}
%
where $0\leq\phi^{\{s\}}\leq 1$ is the mean parous proportion and $\kappa^{\{s\}}>1$ is a concentration parameter (both of which are species-specific).
